---
title: "Additional Exercise 5"
author: "Jonas Moss"
date: "September 20, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


First we load the data:
```{r data, cache = TRUE}
beetle = read.table(file = "http://www.stat.ufl.edu/~aa/glm/data/Beetles2.dat",
                    header = TRUE)
```

# A quadratic logit model
Linear and quadratic logistic regressions can be fitted by doing this.

```{r quadratic logit model, cache = TRUE}

glm(formula = cbind(dead, n - dead) ~ logdose,
    family = binomial(link = logit),
    data = beetle) ->
  linear_logit

glm(formula = cbind(dead, n - dead) ~ logdose + I(logdose^2),
    family = binomial(link = logit),
    data = beetle) ->
  quadratic_logit
```

These can be ploted with the predict function:

```{r quadratic logit model plot, cache = TRUE}

plot(x    = exp(beetle$logdose),
     y    = exp(predict(quadratic_logit))/(1 + exp(predict(quadratic_logit))), 
     type = "b", 
     col  = "blue", 
     bty  = "l",
     ylim = c(0, 1),
     pch  = c(20),
     lwd  = 1.1,
     xlab = "Dose, log scale",
     ylab = "Probability of Death",
     main = "Two logistic models",
     log  = "x")

lines(x    = exp(beetle$logdose),
      y    = exp(predict(linear_logit))/(1 + exp(predict(linear_logit))), 
      type = "b", 
      lwd  = 1.1,
      pch  = c(20),
      col  = "red")

legend(x = "topleft", 
       legend = c("Linear model", "Quadratic model"),
       lty = c(1, 1),
       pch = c(20, 20),
       col = c("red", "blue"),
       bty = "n")
```

We can make use of the `summary` function to assess significance:

```{r quadratic logit model significance, cache = TRUE}

summary(quadratic_logit, corr = TRUE)
```

Hence the correlations between the coefficients are almost as strong as they can
get. This happens because the correlation between `logdose` and `logdose^2` is
very high:

```{r logdose correlation}
cor(beetle$logdose, beetle$logdose^2)
```
 
This correlation can be partly removed by centering `logdose`:

```{r newdose correlation}
newdose = (beetle$logdose - mean(beetle$logdose))
cor(newdose, newdose^2)
```

If we use `newdose`, the correlation between the coefficients estimates nearly
disappears:

```{r quadratic logit model newdose}
glm(formula = cbind(dead, n - dead) ~ newdose + I(newdose^2),
    family = binomial(link = logit),
    data = beetle) ->
 quadratic_logit
summary(quadratic_logit, corr = TRUE)
```


# A Probit Model

We fit both linear and quadratic probit models.

```{r quadratic probit model, cache = TRUE}

glm(formula = cbind(dead, n - dead) ~ logdose,
    family  = binomial(link = probit),
    data    = beetle) ->
  linear_probit

glm(formula = cbind(dead, n - dead) ~ logdose + I(logdose^2),
    family  = binomial(link = probit),
    data    = beetle) ->
  quadratic_probit
```

## Plot

Here is a plot of all the models four models.

```{r quadratic probit model plot, cache = TRUE}

plot(x = beetle$logdose,
     y = exp(predict(quadratic_logit))/(1 + exp(predict(quadratic_logit))), 
     type = "b", 
     col  = "blue", 
     ylim = c(0, 1),
     pch  = c(20),
     xlab = "Log-dose",
     ylab = "Probability of Death",
     main = "Four binary response models.")

lines(x    = beetle$logdose,,
      y    = exp(predict(linear_logit))/(1 + exp(predict(linear_logit))), 
      type = "b", 
      pch  = c(20),
      col  = "red")

lines(x    = beetle$logdose,,
      y    = pnorm(predict(linear_probit)),
      type = "b", 
      lty  = 2,
      pch  = c(20),
      col  = "black")

lines(x    = beetle$logdose,,
      y    = pnorm(predict(quadratic_probit)),
      type = "b", 
      lty  = 2,
      pch  = c(20),
      col  = "green")

legend(x      = "topleft", 
       legend = c("Linear logit", "Quadratic logit", 
                  "Linear probit", "Quadratic probit"),
       lty    = c(1, 1, 2, 2),
       pch    = c(20, 20, 20, 20),
       col    = c("red", "blue", "black", "green"),
       bty    = "n")
```

## Log-likelihoods

The log likelihoods can be computed using he `logLik` function.

```{r likelihooods, cache = TRUE}
models = list(linear_probit, quadratic_probit, linear_logit, quadratic_logit)
unlist(lapply(models, logLik))
```
